---
title: "Importing solar data"
author: "Jason Pither"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message = FALSE, echo = FALSE}
library(readxl)
library(tidyverse)
library(here)
library(stringr)
library(tools)
library(hablar)
library(purrr)
```

## Solar data

First we'll import the solar data downloaded from the [AP Systems website](https://www.apsystemsema.com)  

The raw data files downloaded from the AP Systems portal are in MS Excel format (yuck). They include hourly energy production data for seven consecutive days.  

They include the following:

* "Hour" - the hour of the data (24h), integer value  

* Then seven columns, each holding one day of data, with the column heading being the date (YYYY-MM-DD), with "(kWh)" appended at the end  

* The last row in each data file includes a "total" row.  This will be eliminated.  

Let's first get the list of data files:

```{r get_solar_filenames}
solar.files <- list.files(here::here("rawdata", "solar_data"))
```

We first want to convert all the Excel files into CSV format.

We'll import the data and reformat it to long format, then export to CSV.

First, let's create a vector of filenames we'll use for the output files. The names will include the beginning to end date of the data values.

```{r create_csv_filenames}
solar.output.files <- 1:length(solar.files)

for (i in 1:length(solar.files)) {
  solar.output.files[i] <- solar.files[i] %>% 
    stringr::str_split("Report") %>% 
    purrr::pluck(1, 2) %>% 
    str_trim() %>% 
    str_replace(" to ", "_") %>%
    str_replace(".xls", ".csv")
}
```

Now use those filenames as outputs in this chunk:

```{r loop_import}
for (i in 1:length(solar.files)) {
temp <- readxl::read_xls(here::here("rawdata", "solar_data", 
            solar.files[i]), trim_ws = TRUE) 
# eliminate "total" row
temp <- temp %>% filter(Hour != "Total") %>%
  # convert Hour to integer
  hablar::convert(int(Hour)) %>%
  rename(hour = Hour) %>%
# from wide to long format
  tidyr::pivot_longer(!hour, names_to = "day", values_to = "kWh") %>%
  arrange(day) %>%
    mutate(date = lubridate::ymd_h(paste0(str_sub(day, start = 1, end = 10), hour))) %>%
  select(date, kWh)
# write csv
write_csv(temp, paste(here::here("outdata"), solar.output.files[i], sep = "/"))
  }
```

Get CSV file names

```{r get_csv_filenames}
solar.output.files <- paste(here::here("outdata"), solar.output.files, sep = "/")
```

Import all CSVs into one tibble.

```{r}
solar.data <- readr::read_csv(solar.output.files, id = "file_name")
```

Get rid of file name in table
```{r}
solar.data <- solar.data %>% select(date, kWh)
```

Now create a Day and Hour variable from the "date"
```{r}
solar.data <-solar.data %>%
  mutate(day = lubridate::date(date),
         hour = lubridate::hour(date)) %>%
  select(date, day, hour, kWh)
```

Write csv

```{r}
readr::write_csv(solar.data, here::here("outdata", "solar_data.csv"))
```

## Home energy use data

