---
title: "Importing home energy use data"
author: "Jason Pither"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message = FALSE, echo = FALSE}
library(readxl)
library(tidyverse)
library(here)
library(stringr)
library(tools)
library(hablar)
library(purrr)
library(lubridate)
```

## Home energy use data

First we'll import the Fortis electricity usage data, downloaded from the [Fortis BC website](https://www.fortisbc.com)  

The raw data files downloaded from the Fortis website portal are in CSV format. They include hourly electricity use data for sixty consecutive days.  (Though I may collate these in the end...)

They include the following:

* "Date" - in dd/mm/yyyy format  

* "Time" - in a very strange format of hour (integer) then space then "a.m." or "p.m."  

* "kWh delivered" is a number with precision of 3 decimal places (kWh), representing energy used by the household

* "kWh received" is a number with precision of 3 decimal places (kWh), representing energy supplied back to the grid

Let's first get the list of data files:

```{r get_solar_filenames}
fortis.filenames <- paste(here::here("rawdata", "fortis_data"), 
                  list.files(here::here("rawdata", "fortis_data")), sep = "/")
```

Import CSVs into one tibble.

```{r}
# For some reason the `readr::read_csv` function isn't working properly.. when i try to import multiple CSVs into a single tibble, R aborts.

# So, i'll do a trick that seems to work: reverse the order of files!

# First check if there's more than one file...
if(length(fortis.filenames) > 1) {
fortis.data <- readr::read_csv(fortis.filenames[length(fortis.filenames):1], show_col_types = FALSE)[, -5]
} else { fortis.data <- readr::read_csv(fortis.filenames[1], show_col_types = FALSE) }
```

Having a look at the first few lines, we see a silly format for the "Time" field:
```{r}
fortis.data
```

So, a bit of Googling found this solution on [Stackoverflow](https://stackoverflow.com/questions/72207739/converting-time-to-24-hour-of-day-format-in-r):

```{r}
fortis.data <- fortis.data %>%
  mutate(day = lubridate::dmy(Date),
         hour = hour(parse_date_time(gsub("\\.", "", Time), "%I %p")),
         kw_consumed = `kWh delivered`,
         kw_returned = `kWh received`
         ) %>%
  select(day, hour, kw_returned, kw_consumed) %>%
  arrange(day, hour)
```

Write csv

```{r}
readr::write_csv(fortis.data, here::here("outdata", "fortis_data.csv"))
```

## Billing data

```{r}
billing <- readr::read_csv(here("rawdata", "fortis_billing", "billing_history.csv"))
```

```{r}
billing$read_date <- lubridate::mdy(billing$read_date)
```

```{r secondplot, echo = FALSE, fig.cap = "Figure 1: Historical billing information"}

billing %>%
 # group_by(day) %>%
 # summarise(total = sum(kWh, na.rm = TRUE)) %>%
  ggplot(aes(x = read_date, y = usage)) +
  geom_line(colour = "darkgrey") +
  geom_point(shape = 1) +
  #scale_x_date(date_breaks="4 months") +
  labs(x = "Date", y = "Usage (kWh)") +
 # ylim(0, 90) +
  theme_bw()
```