---
title: "Importing home energy use data"
author: "Jason Pither"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message = FALSE, echo = FALSE}
library(readxl)
library(tidyverse)
library(here)
library(stringr)
library(tools)
library(hablar)
library(purrr)
```

## Home energy use data

First we'll import the Fortis electricity usage data, downloaded from the [Fortis BC website](https://www.fortisbc.com)  

The raw data files downloaded from the Fortis website portal are in CSV format. They include hourly electricity use data for sixty consecutive days.  

They include the following:

* "Date" - in dd/mm/yyyy format  

* "Time" - in a very strange format of hour (integer) then space then "a.m." or "p.m."  

* "kWh delivered" is a number with precision of 3 decimal places (kWh), representing energy used by the household

* "kWh received" is a number with precision of 3 decimal places (kWh), representing energy supplied back to the grid

Let's first get the list of data files:

```{r get_solar_filenames}
fortis.filenames <- paste(here::here("rawdata", "fortis_data"), 
                  list.files(here::here("rawdata", "fortis_data")), sep = "/")
```

Import all CSVs into one tibble.

For some reason the `readr::read_csv` function isn't working properly.. when i try to import multiple CSVs into a single tibble, R aborts.

So, i'll do a trick that seems to work: reverse the order of files!

```{r}
fortis.data <- readr::read_csv(fortis.filenames[length(fortis.filenames):1])
```

Now create a Day and Hour variable from the "date"
```{r}
fortis.data <-fortis.data %>%
  mutate(day = lubridate::date(Date),
         hour = ifelse() %>%
  select(date, day, hour, kWh)
```

Write csv

```{r}
readr::write_csv(solar.data, here::here("outdata", "solar_data.csv"))
```


